### 2018-11-06 Sections/quantum_algorithms.tex
Statement of controls
HLF problem completed

\subsection{Sampling problems}\index{Sampling problems}

@@

\subsection{Quantum advantage with shallow circuits}\index{Shallow circuits}

@@

It's incredibly difficult to compare quantum versus classical polynomial time computation, and currently we have noisy gates in the lab. Given $n$ qubits, circuit depth $d$, and an error rate $\epsilon$, intuitively, we demand the condition,
\begin{align}
nd \ll \frac{1}{\epsilon},
\end{align}
to hold, in order for a quantum algorithm to run successfully. If $d$ is large, then noise dominates and our circuit is classically simulable. On the other hand, if we have large $n$ and small $d$ (constant depth circuits), we may observe a quantum speed up in the near future. Given these factors, we are motivated to examine `shallow circuits'. Now we will see that there is a distinction between constant depth circuits run using classical versus quantum algorithm.

The result is for the following hidden linear function (HLF) problem\index{Hidden linear function problem} \cite{bib:bravyi2018quantum}. Given an  \mbox{$n\times n$} symmetric binary matrix $A$, we specify a quadratic form,
\begin{align}
q(x) = \vec x^T A \vec x \text{(mod 4)}.
\end{align}
The goal is to find a binary vector $\vec z \in \{0,1\}^n$ such that,
\begin{align}
q(x) = 2 \vec z^T \vec x,
\end{align}
for all $\vec x$ in the binary null-space of $A$.

The quantum algorithm that solves the 2D HLF problem is similar to the Bernstein-Vazirani problem \cite{bib:bernstein1997quantum}. Here a quantum circuit $U_q$ acts on the basis state $\vec x \in \{ 0,1\}^n$,
\begin{align}
U_q \ket{x} = i^{q(x)} \ket{x}.
\end{align}
The solution is obtained by measuring the state $\ket{\Psi_q}$ in the computational basis. Here,
\begin{align}
\ket{\Psi_q} &= H^{\otimes n }U_q H^{\otimes n} \ket{0}^{\otimes n} \nonumber \\
&= \frac{1}{\sqrt{2^n}}\sum_{\vec x,\vec z \in \{ 0,1\}^n} i^{q(x)} (-1)^{\vec z^T x}\ket{z},
\end{align}
where $H$ is the Hadamard gate. A schematic of the circuit is given in Fig.~\ref{fig:2DHLF}. Destructive interference between terms in $x$ results a vanishing amplitude \mbox{$\braket{z|\Psi_q}=0$} on basis state $\ket{z}$, which are not solutions of the HLF problem. The circuit $U_q$ can be decomposed into a product of CZ gates and phase-shift gates,
\begin{align}
S_i = \begin{pmatrix}
    1 & 0\\
    0 & -i
    \end{pmatrix}.
\end{align}

\begin{figure}[!htbp]
%\includegraphics[clip=true, width=0.475\textwidth]{2DHLFcircuit}
\caption{\comment{Insert figure}\label{fig:2DHLF} Quantum circuit to solve the 2D HLF problem.}
\end{figure}

@@

Now, comparing the quantum algorithm to a classical one, it has been proven that any probabilistic classical circuit with bounded input that solves the 2D HLF problem with high probability $\epsilon \leq 1/8$ must have depth increasing logarithmically with input size \cite{bib:bravyi2018quantum}. The speed-up here is provided by quantum non-locality\index{Quantum non-locality}.

>>>>

\subsection{Sampling problems}\index{Sampling problems}\label{sec:sampling_problems}

@@

\subsection{Quantum advantage with shallow circuits}\index{Shallow circuits}\label{sec:shallow_circs}

@@

It's incredibly difficult to compare quantum versus classical polynomial time computation, and currently we have only noisy gates in the lab, further complicating the goal of achieving quantum supremacy. Given $n$ qubits, circuit depth $d$, and an error rate $\epsilon$, intuitively, we demand the condition,
\begin{align}
nd \ll \frac{1}{\epsilon},
\end{align}
to hold, in order for a quantum algorithm to run successfully. If $d$ is large, then noise dominates and our circuit is classically simulable. On the other hand, if we have large $n$ and small $d$ (constant depth circuits), we may observe a quantum speed up in the near future. Given these factors, we are motivated to examine `shallow circuits'. Now we will see that there is a distinction between constant depth circuits run using classical versus quantum algorithm.

The result is for the following hidden linear function (HLF) problem\index{Hidden linear function problem} \cite{bib:bravyi2018quantum}. Given an \mbox{$n\times n$} symmetric binary matrix\index{Symmetric binary matrices} $A$, we specify a quadratic form\index{Quadratic form},
\begin{align}
q(x) = \vec x^T A \vec x \, (\mathrm{mod} \,4).
\end{align}
The goal is to find a binary vector $\vec z \in \{0,1\}^n$ such that,
\begin{align}
q(x) = 2 \vec z^T \vec x,
\end{align}
for all $\vec x$ in the binary null-space of $A$.

The quantum algorithm that solves the 2D HLF problem is similar to the Bernstein-Vazirani problem\index{Bernstein-Vazirani problem} \cite{bib:bernstein1997quantum}. The algorithm, shown in Alg.~\ref{alg:shallow_circs}, has similar structure to the IQP sampling problem (Alg.~\ref{alg:IQP_samp}).

\begin{table}[!htbp]
\begin{mdframed}[innertopmargin=3pt, innerbottommargin=3pt, nobreak]
\texttt{
function ShallowQuantumCircuit():
\begin{enumerate}
    \item Prepare the $n$-qubit state,
    \begin{align}
    \ket{\psi_\mathrm{in}} = \ket{0}^{\otimes n}.
    \end{align}
    \item Apply the $n$-qubit Hadamard transform\index{Hadamard transform},
    \begin{align}
    \hat{H}^{\otimes n}.
    \end{align}
    \item Apply the oracle,
	\begin{align}
		U_q \ket{x} = i^{q(x)} \ket{x},
	\end{align}
	for bit-string \mbox{$x\in \{0,1\}^{n}$}, where the oracle is composed of a product of $CZ$ gates\index{Controlled-phase gates} and phase-shifts\index{Phase-shift gates},
	\begin{align}
S_i = \begin{pmatrix}
    1 & 0\\
    0 & -i
    \end{pmatrix}.
\end{align}
	\item Apply another $n$-qubit Hadamard transform,
    \begin{align}
    \hat{H}^{\otimes n}.
    \end{align}
    \item Measure all qubits in the computational basis, yielding measurement outcomes ${\vec z}$.
	\item Repeat protocol $O(\mathrm{poly}(n))$ times.
	\item $\Box$
\end{enumerate}
\begin{align}
\Qcircuit @C=1em @R=1.6em {
    \lstick{\ket{0}} & \gate{\hat{H}} & \multigate{2}{\hat{U}_q} & \gate{\hat{H}} & \meter \\
    \lstick{\vdots} & \gate{\hat{H}} & \ghost{\hat{U}_q} & \gate{\hat{H}} & \meter \\
    \lstick{\ket{0}} & \gate{\hat{H}} & \ghost{\hat{U}_q} & \gate{\hat{H}} & \meter \\
} \nonumber
\end{align}
}
\end{mdframed}
\captionspacealg \caption{Quantum computing using shallow circuits, where circuit depth scales as \mbox{$O(\log d)$}.} \label{alg:shallow_circs}
\end{table}

Here a quantum circuit $U_q$ acts on the computational basis states \mbox{$\vec x \in \{ 0,1\}^n$} via a generalised controlled-phase operation,
\begin{align}
U_q \ket{x} = i^{q(x)} \ket{x},
\end{align}
for some oracle function $q(x)$. The solution is obtained by measuring the state $\ket{\Psi_q}$ in the computational basis. Here,
\begin{align}
\ket{\Psi_q} &= H^{\otimes n }U_q H^{\otimes n} \ket{0}^{\otimes n} \nonumber \\
&= \frac{1}{2^n}\sum_{\vec x,\vec z \in \{ 0,1\}^n} i^{q(x)} (-1)^{\vec z^T x}\ket{z}.
\end{align}

The circuit $U_q$ can be decomposed into a product of CZ gates and phase-shift gates, which all commute, thereby allowing the shallow circuit-depth.

@@

Now, comparing the quantum algorithm to a classical one, it has been proven that any probabilistic classical circuit with bounded input that solves the 2D HLF problem with high probability, $\epsilon \leq 1/8$, must have depth increasing logarithmically with input size \cite{bib:bravyi2018quantum}. The speed-up here is provided by quantum non-locality\index{Quantum non-locality}.
